---
title: "supervised learning in R"
output: html_notebook
---

# Supervised learning - classification and hyperparameter tuning

## Load in data and investigate data

# Load dataset COX-2 Activit

```{r}
data(cox2)
X <- cox2Descr #the descriptors
y <- cox2Class #the categorical outcome ("Active" or "Inactive") based on the $2^2.5$ cutoff
```

# Split the dataset into training (75%) and test (25%) sets

- to get a realistic estimate of our performance after hyperparameter tuning
```{r}
set.seed(123)
inTrain <- createDataPartition(y, p = 0.75, list = FALSE)
X_train <- X[inTrain,]
y_train <-y[inTrain]
X_test <- X[-inTrain,]
y_test <- y[-inTrain]
```

- Check splitting

```{r}
dim(X)
dim(X_train)
dim(X_test)
```

- Identify and exclude zero-variance predictors

```{r}
featVar <- apply(X_train, 2, var)
zerovVarFeat <- featVar == 0
sum(zerovVarFeat)
X_train <- X_train[,featVar>0]
X_test <- X_test[,featVar>0]
```

# Set up preprocessing (scaling/centering) and apply to the training data

```{r}
preProc <- preProcess(X_train, method = c("center", "scale"))
X_train <- predict(preProc, X_train)
```

# Train and evaluate the model

- learning method: random forest

```{r}
set.seed(123)
trControl <- trainControl(method="cv", number=10)  #10-fold cross validation
model_rf <- train(X_train, y_train, 
                  method="rf", 
                  preProcess = c("center","scale"),
                  trainControl=trControl)
```

# Analyze the performance values and feature importance

```{r}
model_rf
importance <- varImp(model_rf)
print(importance)
best_rf <- model_rf$finalModel
best_rf
```

# Apply the same preprocessing to the test data

```{r}
X_test <- predict(preProc, X_test)
```

# Predict on the test set using the final model (best_rf)

```{r}
y_predict <- predict(best_rf, X_test)
confMatrix <- confusionMatrix(y_predict, y_test)
confMatrix
accuracy <- confMatrix$overall['Accuracy']
cat("The accuracy of the random forest model is:", accuracy, "\n")
```